%!TEX root = ../swiatlow_thesis.tex

This thesis and its author owe so much to so many people that it seems like a futile task to try to list them all. I will endeavor to due this list some justice here, and beg foregiveness of those whom I have misplaced from it.

First and foremost, I would like to thank my parents Barbara and Stan: our trips to the Museum of Science and Industry when I was barely walking were where I learned to love science, and set me on the course that has led me hear today. Thank you for that, and for all of the amazing support you have provided over the years. My partner Masha has been the best partner-in-science that I could ask for: I could not imagine having come this far in physics without her patience, encouragement, and comaraderie. 

None of this would have been possible without my advisor Ariel: his devotion, energy, and enthusiasm for jets infected me from the day we met. I could not have asked for a more flexible, kind, helpful advisor. Similarly, my committee members--- Jay Wacker, Blas Cabrera, and Su Dong--- have all contributed enormously to my understanding of particle physics, particle detectors, and experiment in general. My teachers and professors in high school and college are entirely responsible for inspiring my love for physics, symmetries, and hard work: thank you to Neil Koreman, Howard Georgi, and Joao Guimaraes da Costa especially. 

The SLAC group has been a truly wonderful place to work. The post-docs have been a wonderful source of expertise and experience: thank you to Jack, John, Till, Emanuel, and Pascal. My fellow students Katie, Josh, Jim, Dan, Qi, Aviv, and especially Bart and Ben, have made a truly productive and effective team. The number of interesting topics everyone has covered is really remarkable, and I'm honored to have been a part of the gruop.

The best (and worst?) part of ATLAS is having 3000 colleagues--- I have been extraordinarily lucky to work closely with an amazing cast of the best post-docs on the experiment. Thank you to David Miller, David Lopez Mateos, Zach Marshall, Emily Thompson, Lily Asquith, Caterina Doglioni, and Dag Gillberg for all of their mentorship, guidance, and advice. Likewise, more senior advisors such as Peter Loch, Tancredi Carli, and Michael Begel have enormously shaped the way that I approach jets and experiment. My fellow students Larry, Kate, and Tova have not only been wonderful people to work with, but have kept me sane in the darkest depths of deadlines. I also owe an enormous debt also to Gavin Salam, Jesse Thaler, Matt Schwartz, and Jay Wacker: these theorists have been incredibly approachable for a young experimentalist, and I have learned so much and started so many analyses from their direction and advice. The entire community of BOOST--- including our colleages from across the ring--- has been a really wonderful group to discuss techniques and results with.

This list is already getting to the long side, so I will arbitrarily end it here. Thank you to everyone who has answered a grid-support email, spoken up with an idea for me during a meeting, or relaxed with me over a pint after work. I cannot believe I have been so lucky as to travel across the country and the world during the course of a Ph.D., and I cannot believe I have been able to share this experience with so many wonderful people.

